# Reproducible Research - Project 1

## 1. Loading and preprocessing the data

```{r, echo=TRUE}
setwd("C:/Users/Henry/Documents/DataScience/coursera/c05/p01")
rm(list=ls())
data <- read.csv("data/activity.csv")
data$date = as.Date(as.character(data$date), "%Y-%m-%d")
head(data)
str(data)
```

## 2. mean total number of steps taken per day

```{r, echo=TRUE}
#Total number of steps per day
steps.day = tapply(data$steps,data$date, FUN=sum)
steps.day
#Total number of steps over two month period
sum(steps.day, na.rm=TRUE)
#Histogram of steps per day
hist(steps.day, breaks = 10)
#Mean and Median of steps per day
summary(steps.day)
```

The total number of steps per day is 570608, the median is 10765 and The mean is 10766. 

Note that the median and mean are very close, which indicates this distribution behaves like a normal distribution.

## 3. Average daily activity pattern

```{r}
#Calculating Averge number of steps per 5-minute interval
dataClean = data[!is.na(data$steps),]
steps.interval = tapply(dataClean$steps, dataClean$interval, FUN=mean)
minutes = as.integer(names(steps.interval))
#Ploting averge number of steps per 5-minute interval
plot(minutes,steps.interval, type = "l")
#Maximumun Number of steps
max(steps.interval)
#Minute and index with maximun number of steps
which.max(steps.interval)
```

The maximun number of steps is 206.1698 and it occurs at the minute 835.

## 4. Imputing missing values

```{r}
#Calculating summary of the original data to show total number of NA's
summary(data$steps)
#Separate NA values and substiture with the mean for that 5-minute interval
dataNA=data[is.na(data$steps),]
aux = rep(steps.interval, times = 8)
dataNA$steps=aux
#Creating new data set with imputed NA's
data2=rbind(dataClean,dataNA)
#Calculating the total number of steps per day with imputed NA's
steps.day2 = tapply(data2$steps,data2$date, FUN=sum)
#Ploting histograms to compare data with and without imputed NA's
par(mfrow=c(1,2))
hist(steps.day, breaks = 10, main="WITHOUT Imputed NA's", ylim = c(0,25))
hist(steps.day2, breaks = 10, main="WITH Imputed NA's",ylim=c(0,25))
#Summary of data WITHOUT imputed values
summary(steps.day)
#Summary of data WITH imputed values
summary(steps.day2)
```

The strategy for filling the missing values of the dataset was the mean of the 5-Minute Interval. With this strategy, the mean and the median remained the same in both cases. The histogram preserves its shape but it has a higher frequency of observations around the mean.


## 5. Are there differences in activity patterns between weekdays and weekends?

```{r}
#Setting category for Weekdays and Weekends
data2$weekday = weekdays(data2$date)
data2$category = ifelse(data2$weekday %in% c("Saturday","Sunday"),"Weekend","Weekday")
#spliting the data according to category Weekday/Weekend
dataWeekday=data2[data2$category=="Weekday",]
dataWeekend = data2[data2$category=="Weekend",]
#Calculating Averge number of steps per 5-minute interval
steps.weekday = tapply(dataWeekday$steps, dataWeekday$interval, FUN=mean)
steps.weekend = tapply(dataWeekend$steps, dataWeekend$interval, FUN=mean)
minutes.weekday = as.integer(names(steps.weekday))
minutes.weekend = as.integer(names(steps.weekend))
#Ploting averge number of steps per 5-minute interval for weekday
plot(minutes.weekday,steps.weekday, type = "l", ylim = c(0,250))
#Ploting averge number of steps per 5-minute interval for weekend
plot(minutes.weekend,steps.weekend, type = "l", ylim = c(0,250))
```

According to these plots, there is a clear difference in the steps pattern during weekdays and weekends. During Weekdays the activity is less with peaks sparsely distributed during the day. During weekends there is more constant activity throughout the day.